---
title: "Regression"
subtitle: "Part 2b"
author: "Prof. Weldzius"
institute: "Villanova University"
# date: "Slides Updated: `r Sys.Date()`"
output:
  xaringan::moon_reader:
    # self_contained: true
    chakra: libs/remark-latest.min.js
    lib_dir: libs
    css:
      - default
      - css/lexis.css
      - css/lexis-fonts.css
    #seal: false
    nature:
      highlightStyle: github
      highlightLines: true
      countIncrementalSlides: false
      slideNumberFormat: "%current%"
      ratio: "16:9"

---

```{css,echo = F}
.small .remark-code { /*Change made here*/
  font-size: 85% !important;
}
.tiny .remark-code { /*Change made here*/
  font-size: 50% !important;
}
```

```{r,include=F}
set.seed(123)
options(width=60)
knitr::opts_chunk$set(fig.align='center',fig.width=9,fig.height=5,message=F,warning=F)
def.chunk.hook  <- knitr::knit_hooks$get("chunk")
knitr::knit_hooks$set(chunk = function(x, options) {
  x <- def.chunk.hook(x, options)
  ifelse(options$size != "normalsize", paste0("\n \\", options$size,"\n\n", x, "\n\n \\normalsize"), x)
})
```

# Learning Goals

1. Evaluating a regression: Univariate and multivariate visualization of errors

3. Root Mean Squared Error (RMSE)

4. Cross Validation

---

# Remember where we left off last time

```{r,message=F,warning=F,echo=F,include=F}
require(tidyverse)

mv <- read_rds('../data/mv.Rds') %>%
  mutate(gross_log = log(gross),
         budget_log = log(budget))
```

```{r}
m <- lm(gross_log ~ budget_log,data = mv)
summary(m)
```


---

# Evaluation

--

- Every regression line makes mistakes

  - If they didn't, they wouldn't be good at **reducing complexity**!

--

- How bad do ours look?

  - How should we begin to answer this question!?
  
--

- Are there patterns to the mistakes?

  - We **overestimate** gross for movies that cost between $1m and $10m

  - These are the "indies"
  
--

  - We also **underestimate** gross to the "blockbusters"
  
--

- Why?

---

# Understanding Regression Lines

- Regression lines choose $\alpha$ and $\beta$ to minimize mistakes

  - Mistakes (aka "errors" or "residuals") are captured in the $\varepsilon$ term
  
--

  - We can apply the **process** to these!
  
```{r}
# Wrangle data to drop missingness!
mv_analysis <- mv %>% drop_na(gross_log,budget_log)

m <- lm(gross_log ~ budget_log,data = mv_analysis)

mv_analysis$predictions <- predict(m)

mv_analysis$errors <- mv_analysis$gross_log - mv_analysis$predictions

summary(mv_analysis$errors)
```

---

# Univariate Viz of Errors

```{r,message = F}
mv_analysis %>%
  ggplot(aes(x = errors)) + 
  geom_histogram() + 
  labs(x = 'Errors: Gross - Predicted',y = 'Number of movies')
```

---

# Univariate Viz of Errors

- Note that they are on average zero

--

  - Don't feel too proud! Mean 0 error is baked into the method

  - More concerned about **skew**...there is evidence of overestimating
  
--

- Can we do more? **Conditional Analysis**

  - Conditional on the **predictor** (the $X$ variable)
  
---

# Multivariate Viz of Errors

- Ideal is where errors are unrelated to predictor

--

  - This **should** appear as a rectangular cloud of points around zero

```{r,echo=F,message=F,warning=F}
X <- rnorm(1000)
Y <- 2*X + rnorm(1000)

m <- lm(Y~X)

e <- Y - predict(m)

pIdeal <- data.frame(X = X,
           e = e) %>%
ggplot(aes(x = X,y = e)) + 
  geom_point() + 
  geom_hline(yintercept = 0,linetype = 'dashed') + 
  geom_smooth() + 
  labs(x = 'X variable',
       y = 'Errors')
```

```{r,warning=F,message=F}
pIdeal
```


---

# Multivariate Viz of Errors

- This is not the case for us!

```{r,message = F}
mv_analysis %>%
  ggplot(aes(x = budget,y = errors)) + geom_smooth() + geom_point() +
  geom_hline(yintercept = 0,linetype = 'dashed') + scale_x_log10(label = scales::dollar) 
```

---

# Multivariate Viz of Errors

- Evidence of a U-shape &rarr; underpredict low and high budgets, overpredict middle budgets

- Ergo, our model is **not great**!

--

  - Could add additional predictors $X_2$, $X_3$, etc.
  
  - Next lecture!

---

# RMSE

- Univariate / Multivariate visualization of errors is **important**

- But we want to summarize model quality in a simpler way

--

- **RMSE**: summarizes model performance with a *single number*

  - Useful for comparing multiple models to each other


---

# RMSE

- **E**rror ( $\varepsilon$ ): actual outcome ( $Y_i$ ) - predicted outcome ( $\hat{Y}_i$ )

  - The "distance" between the data and the model
  
--

- **S**quared: $\varepsilon^2$

  1. Makes all values positive
  
  2. .blue[Exaggerates] the presence of larger errors
  
--

- **M**ean: average these squared errors

- **R**oot: take their square root (.blue[un-exaggerate])

$$RMSE = \sqrt{\frac{1}{n}\sum_{i = 1}^n(Y_i - \hat{Y}_i)^2}$$

---

# RMSE

- **.red[E]**.red[rror] ( ${\color{red}\varepsilon}$ ): actual outcome ( $Y_i$ ) - predicted outcome ( $\hat{Y}_i$ )

  - The "distance" between the data and the model

- **S**quared: $\varepsilon^2$

  1. Makes all values positive
  
  2. Exaggerates the presence of larger errors

- **M**ean: average these squared errors

- **R**oot: take their square root (un-exaggerate)

$$RMS{\color{red}E} = \sqrt{\frac{1}{n}\sum_{i = 1}^n(\underbrace{Y_i - \hat{Y}_i}_{\varepsilon})^2}$$


---

# RMSE

- **E**rror ( $\varepsilon$ ): actual outcome ( $Y_i$ ) - predicted outcome ( $\hat{Y}_i$ )

  - The "distance" between the data and the model

- **.red[S]**.red[quared:] $\varepsilon^2$

  1. Makes all values positive
  
  2. Exaggerates the presence of larger errors

- **M**ean: average these squared errors

- **R**oot: take their square root (un-exaggerate)

$$RM{\color{red}S}E = \sqrt{\frac{1}{n}\sum_{i = 1}^n(\underbrace{\varepsilon)^2}_{S}}$$

---

# RMSE

- **E**rror ( $\varepsilon$ ): actual outcome ( $Y_i$ ) - predicted outcome ( $\hat{Y}_i$ )

  - The "distance" between the data and the model

- **S**quared: $\varepsilon^2$

  1. Makes all values positive
  
  2. Exaggerates the presence of larger errors

- **.red[M]**.red[ean: average these squared errors]

- **R**oot: take their square root (un-exaggerate)

$$R{\color{red}M}SE = \sqrt{\underbrace{\frac{1}{n}\sum_{i = 1}^n}_{M}(SE)}$$

---

# RMSE

- **E**rror ( $\varepsilon$ ): actual outcome ( $Y_i$ ) - predicted outcome ( $\hat{Y}_i$ )

  - The "distance" between the data and the model

- **S**quared: $\varepsilon^2$

  1. Makes all values positive
  
  2. Exaggerates the presence of larger errors

- **M**ean: average these squared errors

- **.red[R]**.red[oot: take their square root (un-exaggerate)]


$${\color{red}R}MSE = \sqrt{(MSE)}$$

---

# RMSE

- RMSE is a **single measure that summarizes model performance**

--

```{r}
e <- mv_analysis$gross_log - mv_analysis$predictions

se <- e^2

mse <- mean(se)

rmse <- sqrt(mse)
# Or
(rmseBudget <- sqrt(mean(mv_analysis$errors^2)))
```

- Is this good? 


---

# Predicting with uncertainty

- Say we're talking to investors about a new movie that costs $10m

--

  - How do we plug 10m into our model?

```{r, echo=F, results = F}
m <- lm(gross_log ~ budget_log,data = mv)
```

  
```{r}
summary(m)$coefficients
```

- $\hat{Y}_i = \alpha + \beta * X$

--

  - $\alpha = 1.26$ and $\beta = 0.96$
  
--

  - where $\hat{Y}_i$ is predicted gross (log) and $X$ is $10m budget (log)

--

```{r}
pred_gross_log <- 1.26 + 0.96*log(1e7)
```

---

# Predicted Gross

--

- Again, convert back out of logged values with `exp()`

```{r}
scales::dollar(exp(pred_gross_log))
```

--

- Cool! We'll make $8.5m!

--

  - But we know our model isn't perfect
  
--

  - Need to adjust for it's errors via **RMSE**

---

# Incorporating RMSE

- Simple idea: add and subtract RMSE from this prediction

```{r}
pred_gross_log_ub <- 1.26 + 0.96*log(1e7) + rmseBudget
pred_gross_log_lb <- 1.26 + 0.96*log(1e7) - rmseBudget
scales::dollar(exp(c(pred_gross_log_ub,pred_gross_log_lb)))
```

--

- So we'll either make a $56m profit or we'll lose almost $5m?

--

- **CONCLUSION PART 2**: maybe our model isn't very good?


---

# Introducing Cross Validation

- We ran a model on the full data and calculated the RMSE

--

- But this approach risks .red["overfitting"]

--

  - .red[Overfitting] is when we get a model that happens to do well on our specific data, but isn't actually that useful for predicting elsewhere.
  
--

  - "Elsewhere": Other periods, other movies, other datasets
  
--

- .blue[Theory:] Why care about **external validity**?

--

  - What is the point of measuring relationship if they don't generalize?
  
---

# Introducing Cross Validation

- In order to avoid .red[overfitting], we want to "train" our model on one part of the data, and then "test" it on a different part of the data.

--

  - Model "can't see" the test data &rarr; better way to evaluate performance
  
--

- Cross Validation: randomly split our data into a train set and test set

--

  - *Similar to bootstrapping*
  
---

# Introducing Cross Validation (CV)

```{r}
set.seed(1021)
# Create list of row numbers at random
inds <- sample(1:nrow(mv_analysis),
               size = round(nrow(mv_analysis)/2),
               replace = F) 

# Use slice(inds) to get training data
train <- mv_analysis %>% 
  slice(inds)

# Use slice(-inds) to get test data
test <- mv_analysis %>%
  slice(-inds)
```

--

- We now have two datasets of roughly the same number of observations!

---

# CV to Calculate RMSE

--

- We want to estimate a model based on the **test** data

--

- And evaluate RMSE based on the **train** data

```{r}
m2 <- lm(gross_log ~ budget_log,train)

# predict() function on a new dataset
test$preds <- predict(m2,newdata = test)

# Now calculate RMSE on the new dataset
e <- test$gross_log - test$preds
se <- e^2
mse <- mean(se,na.rm=T)
rmse <- sqrt(mse)
rmse
```

---

# CV to Calculate RMSE

- We did worse with CV! This is a *feature*

--

  - We are not being overconfident
  
--

  - We are avoiding "overfitting"
  
--

- Want to do this many times (like bootstrapping)

---

# CV to Calculate RMSE

```{r}
set.seed(123)
bsRes <- NULL
for(i in 1:100) {
  inds <- sample(1:nrow(mv_analysis),
                 size = round(nrow(mv_analysis)/2),
                 replace = F)
  
  train <- mv_analysis %>% slice(inds)
  test <- mv_analysis %>% slice(-inds)
  
  mTrain <- lm(gross_log ~ budget_log,train)
  
  test$preds <- predict(mTrain,newdata = test)
  
  rmse <- sqrt(mean((test$gross_log - test$preds)^2,na.rm=T))
  bsRes <- c(bsRes,rmse)
}

mean(bsRes)
```

---

# CV to Calculate RMSE

```{r}
data.frame(rmseBS = bsRes) %>%
  ggplot(aes(x = rmseBS)) + 
  geom_density() + 
  geom_vline(xintercept = mean(bsRes),linetype = 'dashed')
```


---

# Cross Validation

- In this example, we used a 50-50 split

--

- Often, data scientists prefer an 80-20 split

--

  - **Improves** the model (80% of the data is more to learn from)...
  
  - ...but still **protects** against overfitting
  
```{r}
inds <- sample(1:nrow(mv_analysis),
               size = round(nrow(mv_analysis)*.8), #<<
               replace = F)
```

