---
title: "Regression"
subtitle: "Part 3b"
author: "Prof. Weldzius"
institute: "Villanova University"
# date: "Slides Updated: `r Sys.Date()`"
output:
  xaringan::moon_reader:
    # self_contained: true
    chakra: libs/remark-latest.min.js
    lib_dir: libs
    css:
      - default
      - css/lexis.css
      - css/lexis-fonts.css
    #seal: false
    nature:
      highlightStyle: github
      highlightLines: true
      countIncrementalSlides: false
      slideNumberFormat: "%current%"
      ratio: "16:9"

---

```{css,echo = F}
.small .remark-code { /*Change made here*/
  font-size: 85% !important;
}
.tiny .remark-code { /*Change made here*/
  font-size: 50% !important;
}
```

```{r,include=F}
set.seed(123)
options(width=60)
knitr::opts_chunk$set(fig.align='center',fig.width=9,fig.height=5,message=F,warning=F)
def.chunk.hook  <- knitr::knit_hooks$get("chunk")
knitr::knit_hooks$set(chunk = function(x, options) {
  x <- def.chunk.hook(x, options)
  ifelse(options$size != "normalsize", paste0("\n \\", options$size,"\n\n", x, "\n\n \\normalsize"), x)
})
```

# Agenda

1. Categorical Predictors

---

# Recap of Movie Analysis

```{r,message=F,warning=F}
require(tidyverse)
mv <- read_rds('../data/mv.Rds') %>%
  mutate(gross_log = log(gross),
         budget_log = log(budget))
```

- .blue[Theory]: the more a movie costs, the more it should make

- $X$: log(budget)

- $Y$: log(gross)

---

# Categorical Data

--

- Thus far, only using continuous variables

--

- But we can do regression with categorical data too!

--

- The Bechdel Test: 3 questions of a movie

--

  1. Does it have two women in it?
  2. Who talk to each other?
  3. About something other than a man?
  
```{r}
mv %>%
  count(bechdel_score)
```

---

# Research Question

--

- .blue[Do movies that pass the Bechdel Test make more money?]

--

  - .blue[Theory:] Women are ~50% of the population. Movies that pass the test are more appealing to women.
  
--

  - .blue[Hypothesis:] Movies that pass the test make more money.
  
--

- .red[Wrangling:] Let's turn the `bechdel_score` variable into a binary

```{r}
mv <- mv %>%
  mutate(bechdel_bin = ifelse(bechdel_score == 3,1,0)) %>%
  mutate(bechdel_factor=recode_factor(bechdel_bin,
                                      `1`="Pass",
                                      `0`="Fail",
                                      ))
```

---

# Regression

--

- We can add the binary factor to our regression

```{r}
summary(lm(gross_log ~ bechdel_factor,mv))
```

---

# Regression

- Coefficient is positive

--

- What is the interpretation?

--

  - Movies that fail make more money...
  
--

  - ...than what?
  
--

  - Movies that pass the Bechdel Test
  
--

- Categorical variables are **always interpreted in relation to the hold-out category**!

---

# Regression

- Movies that fail the test make more money!?

--

- **REMEMBER**: Correlation $\neq$ causation

--

  - What might explain this pattern?
  
--

  - Budgets in a sexist Hollywood!
  
--

  - Movies that fail the test get larger budgets
  
--

  - Budgets are positively associated with gross
  
--

- So we want to "control" for budget by adding it to our regression

```{r}
mBechCtrl <- lm(gross_log ~ budget_log + bechdel_factor,mv)
```

---

# Regression

```{r}
summary(mBechCtrl)
```

---

# Regression

- Our hypothesis is supported!

--

- What about non-binary categorical variables?

```{r}
mv %>%
  count(rating)
```

---

# Categorical

- Let's first remove rarely-occurring ratings

```{r}
mvAnalysis <- mv %>%
  filter(!rating %in% c('Approved','TV-14','TV-MA','TV-PG','X'))
```

---

# Categorical

```{r}
summary(lm(gross_log ~ rating,mvAnalysis))
```

---

# Categorical

- Everything makes less money than the hold-out category!

--

  - "G"-rated movies are powered by children
  
--

- What if we wanted to compare to a different reference category?

```{r}
mvAnalysis <- mvAnalysis %>%
  mutate(rating = factor(rating,
                         levels = c('R','PG-13','PG','G','Not Rated')))

mRating2 <- lm(gross_log ~ rating,mvAnalysis)
```

---

# Categorical

```{r}
summary(mRating2)
```

---

# Cross Validation

- This is why `sample_n()` is useful

```{r,warning = F}
set.seed(123)
rmseRes_rating <- NULL
for(i in 1:100) {
  train <- mvAnalysis %>%
    group_by(rating) %>%
    sample_n(size = round(n()*.8),replace = F)
  test <- mvAnalysis %>% anti_join(train)
  
  m <- lm(gross_log ~ rating,train)
  rmseRes_rating <- test %>%
    mutate(preds = predict(m,newdata = test)) %>%
    summarise(rmse = sqrt(mean((gross_log - preds)^2,na.rm=T))) %>%
    bind_rows(rmseRes_rating)
}
rmseRes_rating %>% summarise(rmse = mean(rmse)) #<<
```

---

# Categorical Analysis Practice

- Change the hold-out category to `Not Rated`, run the regression, and cross-validate the results. 

```{r}
# INSERT CODE HERE
```

