---
title: "Classification"
subtitle: "Part 1b"
author: "Prof. Ryan Weldzius"
institute: "Villanova University"
# date: "Slides Updated: `r Sys.Date()`"
output:
  xaringan::moon_reader:
    # self_contained: true
    chakra: libs/remark-latest.min.js
    lib_dir: libs
    css:
      - default
      - css/lexis.css
      - css/lexis-fonts.css
    #seal: false
    nature:
      highlightStyle: github
      highlightLines: true
      countIncrementalSlides: false
      ratio: "16:9"

---

```{css,echo = F}
.small .remark-code { /*Change made here*/
  font-size: 85% !important;
}
.tiny .remark-code { /*Change made here*/
  font-size: 50% !important;
}
```

```{r,include=F}
options(width=60)
knitr::opts_chunk$set(fig.align='center',fig.width=9,fig.height=5,message=F,warning=F)
def.chunk.hook  <- knitr::knit_hooks$get("chunk")
knitr::knit_hooks$set(chunk = function(x, options) {
  x <- def.chunk.hook(x, options)
  ifelse(options$size != "normalsize", paste0("\n \\", options$size,"\n\n", x, "\n\n \\normalsize"), x)
})
```

# Agenda

1. Classification Continued

2. Fortnite gaming

```{r,message=F,warning=F}
require(tidyverse)
require(scales)
fn <- read_rds('../Data/fn_cleaned_final.rds') %>%
  mutate(hits_decile = ntile(hits,n=10)) %>% # Bin hits by decile (10%)
  group_by(hits_decile,mental_state) %>% # Calculate average winning by mental state and accuracy
  mutate(prob_win = mean(won)) %>% # use mutate() instead of summarise() to avoid collapsing the data
  mutate(pred_win = ifelse(prob_win > .5,1,0)) %>% # If the probability is greater than 50-50, predict a win
  ungroup()
```

---

# Recap of Last time

- What is "accuracy"?

  - Proportion "correct" predictions
  
--

- For a binary outcome, "accuracy" has two dimensions

  - Proportion of correct `1`s: **Sensitivity**
  
  - Proportion of correct `0`s: **Specificity**
  
---

# Accuracy

```{r}
(sumTab <- fn %>%
  group_by(won) %>%
  mutate(total_games = n()) %>%
  group_by(won,pred_win,total_games) %>%
  summarise(nGames=n(),.groups = 'drop') %>%
  mutate(prop = nGames / total_games))
```

- Overall accuracy: (`r (sumTab %>% filter(won == 0 & pred_win == 0) %>% pull(nGames))`+`r (sumTab %>% filter(won == 1 & pred_win == 1) %>% pull(nGames))`) / (`r (sumTab %>% filter(won == 0 & pred_win == 0) %>% pull(total_games))`+`r (sumTab %>% filter(won == 1 & pred_win == 1) %>% pull(total_games))`) = `r percent(((sumTab %>% filter(won == 0 & pred_win == 0) %>% pull(nGames)) + (sumTab %>% filter(won == 1 & pred_win == 1) %>% pull(nGames))) / ((sumTab %>% filter(won == 0 & pred_win == 0) %>% pull(total_games)) + (sumTab %>% filter(won == 1 & pred_win == 1) %>% pull(total_games))))`

- But we are doing **great** at predicting losses (`r percent(sumTab %>% filter(won == 0 & pred_win == 0) %>% pull(prop))`)...

- ...and **terribly** at predicting wins (`r percent(sumTab %>% filter(won == 1 & pred_win == 1) %>% pull(prop))`)

---

# Regression

```{r,message = F}
fn %>%
  ggplot(aes(x = damage_to_players,y = won)) + 
  geom_point() + 
  geom_smooth(method = 'lm')
```

---

# Regression

- Binary outcome variable!

--

  - A linear regression is not the best solution
  
--

  - Predictions can exceed support of $Y$

--

- But it can still work! **linear probability model**

```{r}
mLM <- lm(won ~ hits + accuracy + mental_state,fn)
```

---

# Linear Regression

```{r}
require(broom) # broom package makes it easy to read regression output
tidy(mLM) %>% # This would be the same as summary(mLM)
  mutate_at(vars(-term),function(x) round(x,5))
```


---

# Evaluating Predictions

```{r}
mLM <- lm(won ~ hits + accuracy + mental_state + damage_taken + head_shots + gameIdSession,fn)
fn %>%
  mutate(preds = predict(mLM)) %>%
  mutate(predBinary = ifelse(preds > .5,1,0)) %>%
  select(won,predBinary,preds)
```

---

# Evaluating Predictions


```{r}
(sumTab <- fn %>%
  mutate(pred_win = ifelse(predict(mLM) > .5,1,0)) %>%
  group_by(won) %>%
  mutate(total_games = n()) %>%
  group_by(won,pred_win,total_games) %>%
  summarise(nGames=n(),.groups = 'drop') %>%
  mutate(prop = percent(nGames / total_games)) %>%
  ungroup() %>%
  mutate(accuracy = percent(sum((won == pred_win)*nGames) / sum(nGames))))
```

---

# Evaluating Predictions

- Overall accuracy is just the number of correct predictions (either `0` or `1`) out of all possible

--

  - Is `r sumTab$accuracy[1]` good?

--

  - What would the dumbest guess be? Never win! `r percent(1-mean(fn$won))`

--

- Might also want to care about just `1`s

--

  - **Sensitivity**: Predicted wins / actual wins = `r sumTab %>% filter(won == 1 & pred_win == 1) %>% pull(prop)`
  
--

- Also might care about just `0`s

--

  - **Specificity**: Predicted losses / actual losses = `r sumTab %>% filter(won == 0 & pred_win == 0) %>% pull(prop)`
  
---

# Thresholds

- Shifting the threshold for `0` or `1` prediction can matter

--

```{r}
fn %>%
  mutate(pred_win = ifelse(predict(mLM) > .4,1,0)) %>%
  group_by(won) %>%
  mutate(total_games = n()) %>%
  group_by(won,pred_win,total_games) %>%
  summarise(nGames=n(),.groups = 'drop') %>%
  mutate(prop = percent(nGames / total_games)) %>%
  ungroup() %>%
  mutate(accuracy = percent(sum((won == pred_win)*nGames) / sum(nGames)))
```

---

# Thresholds

- Shifting the threshold for `0` or `1` prediction can matter

```{r}
fn %>%
  mutate(pred_win = ifelse(predict(mLM) > .7,1,0)) %>%
  group_by(won) %>%
  mutate(total_games = n()) %>%
  group_by(won,pred_win,total_games) %>%
  summarise(nGames=n(),.groups = 'drop') %>%
  mutate(prop = percent(nGames / total_games)) %>%
  ungroup() %>%
  mutate(accuracy = percent(sum((won == pred_win)*nGames) / sum(nGames)))
```

- Restricting to above 70% means we don't think anyone wins!

---

# Thresholds

- We could keep trying different values until we hit on one that maximizes our accuracy

--

- But this is inefficient! Let's loop it instead!

--


```{r}
toplot <- NULL
for(thresh in seq(0,1,by = .025)) {
  toplot <- fn %>%
  mutate(pred_win = ifelse(predict(mLM) > thresh,1,0)) %>%
  group_by(won) %>%
  mutate(total_games = n()) %>%
  group_by(won,pred_win,total_games) %>%
  summarise(nGames=n(),.groups = 'drop') %>%
  mutate(prop = nGames / total_games) %>%
  ungroup() %>%
  mutate(accuracy = sum((won == pred_win)*nGames) / sum(nGames)) %>%
  mutate(threshold = thresh) %>%
    bind_rows(toplot)
}
```

---

# Thresholds

- We might only care about accuracy by itself (although this is a bit naive)

.small[
```{r}
toplot %>%
  select(accuracy,threshold) %>%
  distinct() %>%
  ggplot(aes(x = threshold,y = accuracy)) + 
  geom_line()
```
]

---

# Thresholds

.small[
```{r}
toplot %>%
  mutate(metric = ifelse(won == 1 & pred_win == 1,'Sensitivity',
                         ifelse(won == 0 & pred_win == 0,'Specificity',NA))) %>%
  drop_na(metric) %>%
  ggplot(aes(x = threshold,y = prop,color = metric)) + 
  geom_line()
```
]


---

# ROC Curve

- Receiver-Operator Characteristic (ROC) Curve

--

- Commonly used to evaluate classification methods

--

  - X-axis: 1-specificity

  - Y-axis: sensitivity

--

```{r}
p <- toplot %>%
  mutate(metric = ifelse(won == 1 & pred_win == 1,'Sensitivity',
                         ifelse(won == 0 & pred_win == 0,'Specificity',NA))) %>%
  drop_na(metric) %>%
  select(prop,metric,threshold) %>%
  spread(metric,prop) %>%
  arrange(desc(Specificity),Sensitivity) %>%
  ggplot(aes(x = 1-Specificity,y = Sensitivity)) + 
  geom_line() + 
  xlim(c(0,1)) + ylim(c(0,1)) + 
  geom_abline(slope = 1,intercept = 0,linetype = 'dotted') + 
  ggridges::theme_ridges()
```

---

# ROC Curve

```{r}
p
```

--

- Better models have high levels of sensitivity **and** specificity at every threshold

---

# AUC Measure

- Area Under the Curve (AUC)

--

  - A single number summarizing classification performance
  
--

```{r,message=F}
require(tidymodels)
roc_auc(data = fn %>%
  mutate(pred_win = predict(mLM),
         truth = factor(won,levels = c('1','0'))) %>%
  select(truth,pred_win),truth,pred_win)
```

---

# AUC

- What is a "good" AUC?

--

  - We know it is bounded between 0 (i.e., it predicts everything **perfectly wrong**) and 1 (i.e., it predicts everything **perfectly correct**)
  
  - But typically we don't see AUC values less than 0.5 (why is this?)
  
--

- AUC can be interpreted like numeric grades at Villanova (and for this class)

  - 0.95+ is amazing
  
  - 0.9 - 0.95 is very good
  
  - 0.8-range is B-tier
  
  - 0.7-range is C-tier
  
  - 0.6-range is really bad
  
  - AUC values less than 0.6 are failing

---

# Party time!

- Adding more variables / trying different combinations

--

- **Workflow**

--

  1. Train models
  
  2. Predict models
  
  3. Evaluate models
  
---

# Train models

```{r}
m1 <- lm(won ~ hits,fn)
m2 <- lm(won ~ hits + head_shots,fn)
m3 <- lm(won ~ hits + accuracy + head_shots,fn)
m4 <- lm(won ~ hits + accuracy + head_shots + mental_state,fn)
m5 <- lm(won ~ hits + accuracy + head_shots + mental_state + distance_traveled,fn)
m6 <- lm(won ~ hits + accuracy + mental_state + head_shots + distance_traveled + gameIdSession,fn)
```

---

# Predict models

```{r}
toEval <- fn %>%
  mutate(m1Preds = predict(m1),
         m2Preds = predict(m2),
         m3Preds = predict(m3),
         m4Preds = predict(m4),
         m5Preds = predict(m5),
         m6Preds = predict(m6),
         truth = factor(won,levels = c('1','0')))
```

---

# Evaluate models

```{r}
rocRes <- NULL
for(model in 1:6) {
  rocRes <- roc_auc(toEval,truth,paste0('m',model,'Preds')) %>%
    mutate(model = paste0('Model ',model)) %>%
    bind_rows(rocRes)
}
```

---

# Evaluate models

```{r}
rocRes %>%
  ggplot(aes(x = .estimate,y = reorder(model,.estimate))) + 
  geom_bar(stat = 'identity') + 
  ggridges::theme_ridges() + labs(x = 'AUC',y = 'Regression Model')
```

---

# OVERFITTING

- Cross validation to the rescue!

.tiny[
```{r}
set.seed(123)
cvRes <- NULL
for(i in 1:100) {
  # Cross validation prep
  inds <- sample(1:nrow(fn),size = round(nrow(fn)*.8),replace = F)
  train <- fn %>% slice(inds)
  test <- fn %>% slice(-inds)

  # Training models
  m1 <- lm(won ~ hits,train)
  m2 <- lm(won ~ hits + head_shots,train)
  m3 <- lm(won ~ hits + accuracy + head_shots,train)
  m4 <- lm(won ~ hits + accuracy + head_shots + mental_state,train)
  m5 <- lm(won ~ hits + accuracy + head_shots + mental_state + distance_traveled,train)
  m6 <- lm(won ~ hits + accuracy + mental_state + head_shots + distance_traveled + gameIdSession,train)

  # Predicting models
  toEval <- test %>%
    mutate(m1Preds = predict(m1,newdata = test),
           m2Preds = predict(m2,newdata = test),
           m3Preds = predict(m3,newdata = test),
           m4Preds = predict(m4,newdata = test),
           m5Preds = predict(m5,newdata = test),
           m6Preds = predict(m6,newdata = test),
           truth = factor(won,levels = c('1','0')))

  # Evaluating models
  rocResBS <- NULL
  for(model in 1:6) {
    rocResBS <- roc_auc(toEval,truth,paste0('m',model,'Preds')) %>%
      mutate(model = as.character(get(paste0('m',model))$call$formula)[3]) %>%
      bind_rows(rocResBS)
  }
  cvRes <- rocResBS %>%
    mutate(bsInd = i) %>%
    bind_rows(cvRes)
}
```
]
---

# Cross Validation AUC

```{r}
cvRes %>%
  ggplot(aes(x = .estimate,y = factor(reorder(model,.estimate)))) + 
  geom_boxplot() + labs(x = 'Distribution of AUC',y = 'Specification')
```

---

# Conclusion

- Classification is just a type of prediction

--

  - We used linear regression
  
--

  - But there are **much** fancier algorithms out there
  
--

- Next class:

  - A *slightly* fancier algorithm: logistic regression
  
  - How to use the models to achieve the team's goals
  

