---
title: "Clustering"
subtitle: "Part 1b"
author: "Prof. Ryan Weldzius"
institute: "Villanova University"
# date: "Slides Updated: `r Sys.Date()`"
output:
  xaringan::moon_reader:
    # self_contained: true
    chakra: libs/remark-latest.min.js
    lib_dir: libs
    css:
      - default
      - css/lexis.css
      - css/lexis-fonts.css
    #seal: false
    nature:
      highlightStyle: github
      highlightLines: true
      countIncrementalSlides: false
      ratio: "16:9"

---

```{css,echo = F}
.small .remark-code { /*Change made here*/
  font-size: 85% !important;
}
.tiny .remark-code { /*Change made here*/
  font-size: 50% !important;
}
```

```{r,include=F}
set.seed(123)
options(width=60)
knitr::opts_chunk$set(fig.align='center',fig.width=9,fig.height=5,message=F,warning=F)
def.chunk.hook  <- knitr::knit_hooks$get("chunk")
knitr::knit_hooks$set(chunk = function(x, options) {
  x <- def.chunk.hook(x, options)
  ifelse(options$size != "normalsize", paste0("\n \\", options$size,"\n\n", x, "\n\n \\normalsize"), x)
})
```

# Agenda

1. Application of k-means clustering

---

# Clustering to Learn about MCs

```{r, message=FALSE}
library(tidyverse)
dat <- read_csv('../Data/H097_members.csv')
glimpse(dat)
```

---

# DW-NOMINATE

- DW-NOMINATE is a measure of how frequently different legislators vote together

--

- Often interpreted as "ideology"

--

- Two-dimensions:

--

  1. Standard left-right ideology (size of gov, redistribution, etc.)
  
  2. Second dimension changes, but typically **salient social issues**
  
--

- Can $k$-means clustering help us learn about legislators?

---

# 97th Congress (1981-1983)

```{r,message = F,warning = F}
require(scales)
library(plotly)
gg <- dat %>%
  ggplot(aes(x = nominate_dim1,y = nominate_dim2,
             text = bioname)) + 
  geom_point() + 
  labs(x = 'DW-Nominate Dimension 1',
       y = 'DW-Nominate Dimension 2')
```

---

# 97th Congress (1981-1983)

```{r,warning = F,echo = F}
ggplotly(gg,tooltip = "text")
```

---

# Intuition Check

--

- Can we see some clusters?

  - What do we think these are?
  
--

- Let's try estimating $k$-means!

--

- Function `kmeans(x,centers,iter.max,nstart)`

--

  - `x` is the data (only select the columns of interest!)
  
  - `centers` is the number of centroids
  
  - `iter.max` maximum amount of "steps"
  
  - `nstart` how many times to re-estimate
  
---

# Clustering on Ideology

- First, some light wrangling (convert numeric party code to character)

```{r}
datClust <- dat %>% 
  mutate(party = ifelse(party_code == 200,'R',
                        ifelse(party_code == 100,'D','I'))) %>%
  mutate(nameParty = paste0(bioname,' (',party,')')) %>%
  select(nominate_dim1,nominate_dim2,nameParty) %>% drop_na()
```

---

# Clustering on Ideology

- Second, estimate `kmeans()` function

```{r}
m <- kmeans(x = datClust %>% select(nominate_dim1,nominate_dim2),
            centers = 2)

m
```

---

# Clustering on Ideology

- Easier to see output with the help of `tidymodels` package

```{r,message=F,warning=F}
require(tidymodels)
tidy(m)
```

--

- First two columns are the **locations** of the centroids

- `size` is the number of observations associated with each group

- `withinss` is the **errors** each centroid makes

---

# Clustering on Ideology

- Third, plot points and color by cluster

```{r}
ggClust <- datClust %>%
  mutate(cluster = m$cluster) %>% # Add cluster to data
  ggplot(aes(x = nominate_dim1, y = nominate_dim2, 
             color = factor(cluster),
             text=nameParty)) +
  geom_point() +
  labs(x = 'DW-Nominate Dimension 1',
       y = 'DW-Nominate Dimension 2',
       title="97th Congress Ideology",
       color = "Cluster")
```

---

# Clustering on Ideology

```{r}
ggplotly(ggClust,tooltip = 'text')
```


---

# More Clusters

```{r}
m <- kmeans(x = datClust %>% select(nominate_dim1,nominate_dim2),
            centers = 3)
```
```{r,echo = F}
ggClust <- datClust %>%
  mutate(cluster = m$cluster) %>%
  ggplot(aes(x = nominate_dim1, y = nominate_dim2, color = factor(cluster),
             text=nameParty)) +
  geom_point() +
  labs(x = 'DW-Nominate Dimension 1',
       y = 'DW-Nominate Dimension 2',
       title="97th Congress Ideology",
       subtitle = '# Clusters = 3',
       color = "Cluster")
ggplotly(ggClust,tooltip = 'text')
```


---

# More Clusters

```{r}
m <- kmeans(x = datClust %>% select(nominate_dim1,nominate_dim2),
            centers = 4)
```
```{r,echo = F}
ggClust <- datClust %>%
  mutate(cluster = m$cluster) %>%
  ggplot(aes(x = nominate_dim1, y = nominate_dim2, color = factor(cluster),
             text=nameParty)) +
  geom_point() +
  labs(x = 'DW-Nominate Dimension 1',
       y = 'DW-Nominate Dimension 2',
       title="97th Congress Ideology",
       subtitle = '# Clusters = 4',
       color = "Cluster")
ggplotly(ggClust,tooltip = 'text')
```


---

# More Clusters

```{r}
m <- kmeans(x = datClust %>% select(nominate_dim1,nominate_dim2),
            centers = 5)
```
```{r,echo = F}
ggClust <- datClust %>%
  mutate(cluster = m$cluster) %>%
  ggplot(aes(x = nominate_dim1, y = nominate_dim2, color = factor(cluster),
             text=nameParty)) +
  geom_point() +
  labs(x = 'DW-Nominate Dimension 1',
       y = 'DW-Nominate Dimension 2',
       title="97th Congress Ideology",
       subtitle = '# Clusters = 5',
       color = "Cluster")
ggplotly(ggClust,tooltip = 'text')
```

---

# How many clusters?

--

- Recall from regression that we are interested in **errors**

--

- What are "errors" in the context of clustering?

--

```{r,echo = F}
m <- datClust %>% select(nominate_dim1,nominate_dim2) %>%
  kmeans(centers = 2)

cents <- data.frame(m$centers)

toplot <- datClust %>%
  mutate(clust = m$cluster) 

toplot %>%
  ggplot(aes(x = nominate_dim1,y = nominate_dim2,color = factor(clust))) + 
  geom_point() + 
  geom_point(data = cents,aes(x = nominate_dim1,y = nominate_dim2),
             inherit.aes = F,shape = '+',size = 10) + 
  labs(x = 'DW-Nominate Dimension 1',
       y = 'DW-Nominate Dimension 2',
       title="97th Congress Ideology",
       color = "Cluster") + 
  theme(legend.position = 'none')
```

---

# How many clusters?

- Recall from regression that we are interested in **errors**

- What are "errors" in the context of clustering?

```{r,echo = F}
toplot %>%
  ggplot(aes(x = nominate_dim1,y = nominate_dim2,color = factor(clust))) + 
  geom_point() + 
  geom_segment(data = toplot %>%
                 filter(clust == 1),
               aes(x = nominate_dim1,y = nominate_dim2),xend = cents$nominate_dim1[1],yend = cents$nominate_dim2[1]) +
  geom_segment(data = toplot %>%
                 filter(clust == 2),
               aes(x = nominate_dim1,y = nominate_dim2),xend = cents$nominate_dim1[2],yend = cents$nominate_dim2[2]) +
  geom_point(data = cents,aes(x = nominate_dim1,y = nominate_dim2),
             inherit.aes = F,shape = '+',size = 10) + 
  labs(x = 'DW-Nominate Dimension 1',
       y = 'DW-Nominate Dimension 2',
       title="97th Congress Ideology",
       color = "Cluster") + 
  theme(legend.position = 'none')
```


---

# How many clusters?

- Recall from regression that we are interested in **errors**

- What are "errors" in the context of clustering?

- Just the sum of each observation's distance from its centroid!

--

  - Within Sum of Squares (**WSS**)

```{r,message=F,warning=F}
tidy(m)
```

---

# So...how many?!

- Want to choose a number of clusters that reduces the total **WSS**

--

```{r}
m.cluster <- datClust %>%
  select(-nameParty) %>% # Same as selecting two dimensions
  kmeans(centers = 3)
```

```{r,echo = F,warning=F,message=F}
cents <- data.frame(m.cluster$centers) %>%
  mutate(clust = row_number()) %>%
  rename(xend = nominate_dim1,yend = nominate_dim2)

cents$wss <- m.cluster$withinss

toplot <- datClust %>%
  as_tibble() %>%
  mutate(clust = m.cluster$cluster) %>%
  left_join(cents)

toplot %>%
  ggplot(aes(x = nominate_dim1,y = nominate_dim2,color = factor(clust))) + 
  geom_point() + 
  geom_segment(aes(x = nominate_dim1,y = nominate_dim2,xend = xend,yend = yend)) + 
  geom_point(data = cents,aes(x = xend,y = yend),
             inherit.aes = F,shape = '+',size = 10) + 
  geom_label(data = cents,aes(x = xend,y = yend,label = round(wss,1)),inherit.aes = F,hjust = 0,vjust = 0,nudge_x = .1,nudge_y = .1) + 
  annotate(geom = 'label',x = 0,y = 1,label = paste0('Tot WSS = ',round(m.cluster$tot.withinss,1))) + 
  labs(x = 'DW-Nominate Dimension 1',
       y = 'DW-Nominate Dimension 2',
       title="97th Congress Ideology",
       color = "Cluster") + 
  theme(legend.position = 'none')
```

---

# So...how many?!

- Want to choose a number of clusters that reduces the total **WSS**

```{r}
m.cluster <- datClust %>%
  select(-nameParty) %>%
  kmeans(centers = 4)
```

```{r,echo = F,warning=F,message=F}
cents <- data.frame(m.cluster$centers) %>%
  mutate(clust = row_number()) %>%
  rename(xend = nominate_dim1,yend = nominate_dim2)

cents$wss <- m.cluster$withinss

toplot <- datClust %>%
  as_tibble() %>%
  mutate(clust = m.cluster$cluster) %>%
  left_join(cents)

toplot %>%
  ggplot(aes(x = nominate_dim1,y = nominate_dim2,color = factor(clust))) + 
  geom_point() + 
  geom_segment(aes(x = nominate_dim1,y = nominate_dim2,xend = xend,yend = yend)) + 
  geom_point(data = cents,aes(x = xend,y = yend),
             inherit.aes = F,shape = '+',size = 10) + 
  geom_label(data = cents,aes(x = xend,y = yend,label = round(wss,1)),inherit.aes = F,hjust = 0,vjust = 0,nudge_x = .1,nudge_y = .1) + 
  annotate(geom = 'label',x = 0,y = 1,label = paste0('Tot WSS = ',round(m.cluster$tot.withinss,1))) + 
  labs(x = 'DW-Nominate Dimension 1',
       y = 'DW-Nominate Dimension 2',
       title="97th Congress Ideology",
       color = "Cluster") + 
  theme(legend.position = 'none')
```

---

# So...how many?!

- Want to choose a number of clusters that reduces the total **WSS**

```{r}
m.cluster <- datClust %>%
  select(-nameParty) %>%
  kmeans(centers = 5)
```

```{r,echo = F,warning=F,message=F}
cents <- data.frame(m.cluster$centers) %>%
  mutate(clust = row_number()) %>%
  rename(xend = nominate_dim1,yend = nominate_dim2)

cents$wss <- m.cluster$withinss

toplot <- datClust %>%
  as_tibble() %>%
  mutate(clust = m.cluster$cluster) %>%
  left_join(cents)

toplot %>%
  ggplot(aes(x = nominate_dim1,y = nominate_dim2,color = factor(clust))) + 
  geom_point() + 
  geom_segment(aes(x = nominate_dim1,y = nominate_dim2,xend = xend,yend = yend)) + 
  geom_point(data = cents,aes(x = xend,y = yend),
             inherit.aes = F,shape = '+',size = 10) + 
  geom_label(data = cents,aes(x = xend,y = yend,label = round(wss,1)),inherit.aes = F,hjust = 0,vjust = 0,nudge_x = .1,nudge_y = .1) + 
  annotate(geom = 'label',x = 0,y = 1,label = paste0('Tot WSS = ',round(m.cluster$tot.withinss,1))) + 
  labs(x = 'DW-Nominate Dimension 1',
       y = 'DW-Nominate Dimension 2',
       title="97th Congress Ideology",
       color = "Cluster") + 
  theme(legend.position = 'none')
```

---

# So...how many?!

- Want to choose a number of clusters that reduces the total **WSS**

```{r}
m.cluster <- datClust %>%
  select(-nameParty) %>%
  kmeans(centers = 10)
```

```{r,echo = F,warning=F,message=F}
cents <- data.frame(m.cluster$centers) %>%
  mutate(clust = row_number()) %>%
  rename(xend = nominate_dim1,yend = nominate_dim2)

cents$wss <- m.cluster$withinss

toplot <- datClust %>%
  as_tibble() %>%
  mutate(clust = m.cluster$cluster) %>%
  left_join(cents)

toplot %>%
  ggplot(aes(x = nominate_dim1,y = nominate_dim2,color = factor(clust))) + 
  geom_point() + 
  geom_segment(aes(x = nominate_dim1,y = nominate_dim2,xend = xend,yend = yend)) + 
  geom_point(data = cents,aes(x = xend,y = yend),
             inherit.aes = F,shape = '+',size = 10) + 
  geom_label(data = cents,aes(x = xend,y = yend,label = round(wss,1)),inherit.aes = F,hjust = 0,vjust = 0,nudge_x = .1,nudge_y = .1) + 
  annotate(geom = 'label',x = 0,y = 1,label = paste0('Tot WSS = ',round(m.cluster$tot.withinss,1))) + 
  labs(x = 'DW-Nominate Dimension 1',
       y = 'DW-Nominate Dimension 2',
       title="97th Congress Ideology",
       color = "Cluster") + 
  theme(legend.position = 'none')
```

---

# So...how many?!

- Want to choose a number of clusters that reduces the total **WSS**

```{r}
m.cluster <- datClust %>%
  select(-nameParty) %>%
  kmeans(centers = 30)
```

```{r,echo = F,warning=F,message=F}
cents <- data.frame(m.cluster$centers) %>%
  mutate(clust = row_number()) %>%
  rename(xend = nominate_dim1,yend = nominate_dim2)

cents$wss <- m.cluster$withinss

toplot <- datClust %>%
  as_tibble() %>%
  mutate(clust = m.cluster$cluster) %>%
  left_join(cents)

toplot %>%
  ggplot(aes(x = nominate_dim1,y = nominate_dim2,color = factor(clust))) + 
  geom_point() + 
  geom_segment(aes(x = nominate_dim1,y = nominate_dim2,xend = xend,yend = yend)) + 
  geom_point(data = cents,aes(x = xend,y = yend),
             inherit.aes = F,shape = '+',size = 10) + 
  geom_label(data = cents,aes(x = xend,y = yend,label = round(wss,1)),inherit.aes = F,hjust = 0,vjust = 0,nudge_x = .1,nudge_y = .1) + 
  annotate(geom = 'label',x = 0,y = 1,label = paste0('Tot WSS = ',round(m.cluster$tot.withinss,1))) + 
  labs(x = 'DW-Nominate Dimension 1',
       y = 'DW-Nominate Dimension 2',
       title="97th Congress Ideology",
       color = "Cluster") + 
  theme(legend.position = 'none')
```

---

# So...how many?!

- But there's a trade-off!

--

  - Accuracy versus **parsimony**
  
--

- Simple rule: look for the "elbow"

```{r}
totWSS <- NULL
for(k in 1:10) {
  m.cluster <- datClust %>%
    select(-nameParty) %>%
    kmeans(centers = k,nstart = 25)
  totWSS <- data.frame(totWSS = m.cluster$tot.withinss,
             k = k) %>%
    bind_rows(totWSS)
}
```

---

# Looking for the "elbow"

```{r}
totWSS %>%
  ggplot(aes(x = k,y = totWSS)) + 
  geom_line() + geom_point() + 
  labs(x = 'Number of Clusters',y = 'Total WSS') + 
  scale_x_continuous(breaks = 1:10)
```


---

# Clustering on Ideology

- Note that we need to `set.seed()`! Centroid starting points are **random**!

```{r}
m <- kmeans(x = datClust %>% select(nominate_dim1,nominate_dim2),
                    centers = 2)
```
```{r,echo = F}
ggClust <- datClust %>%
  mutate(cluster = m$cluster) %>%
  ggplot(aes(x = nominate_dim1, y = nominate_dim2, color = factor(cluster),
             text=nameParty)) +
  geom_point() +
  labs(x = 'DW-Nominate Dimension 1',
       y = 'DW-Nominate Dimension 2',
       title="97th Congress Ideology",
       color = "Cluster")
ggplotly(ggClust,tooltip = 'text')
```

---

# Clustering on Ideology

- Note that we need to `set.seed()`! Centroid starting points are **random**!

```{r}
m <- kmeans(x = datClust %>% select(nominate_dim1,nominate_dim2),
                    centers = 2)
```
```{r,echo = F}
ggClust <- datClust %>%
  mutate(cluster = m$cluster) %>%
  ggplot(aes(x = nominate_dim1, y = nominate_dim2, color = factor(cluster),
             text=nameParty)) +
  geom_point() +
  labs(x = 'DW-Nominate Dimension 1',
       y = 'DW-Nominate Dimension 2',
       title="97th Congress Ideology",
       color = "Cluster")
ggplotly(ggClust,tooltip = 'text')
```

---

# Clustering on Ideology

- Note that we need to `set.seed()`! Centroid starting points are **random**!

```{r}
m <- kmeans(x = datClust %>% select(nominate_dim1,nominate_dim2),
                    centers = 2)
```
```{r,echo = F}
ggClust <- datClust %>%
  mutate(cluster = m$cluster) %>%
  ggplot(aes(x = nominate_dim1, y = nominate_dim2, color = factor(cluster),
             text=nameParty)) +
  geom_point() +
  labs(x = 'DW-Nominate Dimension 1',
       y = 'DW-Nominate Dimension 2',
       title="97th Congress Ideology",
       color = "Cluster")
ggplotly(ggClust,tooltip = 'text')
```

---

# Clustering on Ideology

- Note that we need to `set.seed()`! Centroid starting points are **random**!

```{r}
m <- kmeans(x = datClust %>% select(nominate_dim1,nominate_dim2),
                    centers = 2)
```
```{r,echo = F}
ggClust <- datClust %>%
  mutate(cluster = m$cluster) %>%
  ggplot(aes(x = nominate_dim1, y = nominate_dim2, color = factor(cluster),
             text=nameParty)) +
  geom_point() +
  labs(x = 'DW-Nominate Dimension 1',
       y = 'DW-Nominate Dimension 2',
       title="97th Congress Ideology",
       color = "Cluster")
ggplotly(ggClust,tooltip = 'text')
```

---

# Clustering Randomness

- Can overcome with `nstart`

--

  - Attempts multiple initial centroids and chooses the "best"
  
```{r}
set.seed(42)
c1 <- kmeans(datClust %>% select(-nameParty),centers = 2,nstart = 25)

set.seed(123)
c2 <- kmeans(datClust %>% select(-nameParty),centers = 2,nstart = 25)

table(c1$cluster,c2$cluster)

table(c1$cluster)
table(c2$cluster)
```




---

# Is Polarization Increasing?

- Compare 97th Congress (1981-1983) to 117th Congress (2021-2023)

```{r,warning=F,message=F}
dat <- read_csv('../Data/H117_members.csv')
datClust <- dat %>% 
  mutate(party = ifelse(party_code == 200,'R',
                        ifelse(party_code == 100,'D','I'))) %>%
  mutate(nameParty = paste0(bioname,' (',party,')')) %>% # combine name + party
  select(nominate_dim1,nominate_dim2,nameParty) %>% drop_na()
```

- Check for the "elbow"

```{r}
totWSS <- NULL
for(k in 1:10) {
  m.cluster <- datClust %>%
    select(-nameParty) %>% kmeans(centers = k,nstart = 25)
  totWSS <- data.frame(totWSS = m.cluster$tot.withinss,k = k) %>%
    bind_rows(totWSS)
}
```

---

# Check for the "elbow"

```{r}
totWSS %>%
  ggplot(aes(x = k,y = totWSS)) + 
  geom_line() + geom_point() + 
  labs(x = 'Number of Clusters',y = 'Total WSS') + 
  scale_x_continuous(breaks = 1:10)
```

---

# Growing polarization?

```{r}
m <- kmeans(x = datClust %>% select(nominate_dim1,nominate_dim2),
            centers = 2)
```
```{r,echo = F}
ggClust <- datClust %>%
  mutate(cluster = m$cluster) %>%
  ggplot(aes(x = nominate_dim1, y = nominate_dim2, color = factor(cluster),
             text=nameParty)) +
  geom_point() +
  labs(x = 'DW-Nominate Dimension 1',
       y = 'DW-Nominate Dimension 2',
       title="117th Congress Ideology",
       color = "Cluster")
ggplotly(ggClust,tooltip = 'text')
```


---

# Conclusion

--

- Clustering is part of a third camp of data science

--

  1. Theory Testing
  
  2. Prediction
  
  3. **Learning**
  
