---
title: "Uncertainty Part 2b"
subtitle: "Sports Analytic Mania: Bootstrapping!"
author: "Prof. Weldzius"
institute: "Villanova University"
# date: "Slides Updated: `r Sys.Date()`"
output:
  xaringan::moon_reader:
    # self_contained: true
    chakra: libs/remark-latest.min.js
    lib_dir: libs
    css:
      - default
      - css/lexis.css
      - css/lexis-fonts.css
    #seal: false
    nature:
      highlightStyle: github
      highlightLines: true
      countIncrementalSlides: false
      slideNumberFormat: "%current%"
      ratio: "16:9"

---

```{css,echo = F}
.small .remark-code { /*Change made here*/
  font-size: 85% !important;
}
.tiny .remark-code { /*Change made here*/
  font-size: 50% !important;
}
```

```{r,include=F}
set.seed(123)
options(width=60)
knitr::opts_chunk$set(fig.align='center',fig.width=9,fig.height=5,message=F,warning=F)
def.chunk.hook  <- knitr::knit_hooks$get("chunk")
knitr::knit_hooks$set(chunk = function(x, options) {
  x <- def.chunk.hook(x, options)
  ifelse(options$size != "normalsize", paste0("\n \\", options$size,"\n\n", x, "\n\n \\normalsize"), x)
})
```

# Agenda

1. Uncertainty

2. More NBA data

3. Bootstrap Sampling!

---

# Sports Analytics Continued

- Load the `game_summary.Rds` data

```{r,message=F,warning=F}
require(tidyverse)
gms <- read_rds('../data/game_summary.Rds')
gms
```

---

# Remember the .blue[Science]

--

- What predicts winning?

  - Points? (more is better)
  - Turnovers? (less is better)
  - Rebounds? (more is better)
  
--

- On average, winning teams have ~1 fewer turnover than losing teams

- FSNoR: is this *always* the case?

- Use **bootstrap sampling** to see!
  

---

# Looping

```{r}
set.seed(123)
bs_tov <- NULL
for(i in 1:1000) {
  bs_tov <- gms %>%
    sample_n(size = 100,replace = T) %>%
    group_by(isWin) %>%
    summarise(avgTO = mean(tov)) %>%
    bind_rows(bs_tov)
}
bs_tov %>% head()
```

---

# Bootstrapped Estimates vs Data

```{r}
bs_tov %>%
  group_by(isWin) %>%
  summarise(bs_est = mean(avgTO))

gms %>%
  group_by(isWin) %>%
  summarise(data_est = mean(tov))
```

---

# Bootstrapped Estimates vs Data

--

- They're identical!

--

  - In .blue[theory], bootstrapped samples converge on true values
  
--

  - ...where "true" is the full data
  
--

- So then why bother with bootstrapping?

--

- **Uncertainty!**

---

# Plot Distributions of Bootstraps

```{r,fig.width=9,fig.height=5}
bs_tov %>%
  ggplot(aes(x = avgTO,fill = isWin)) + 
  geom_density(alpha = .3)
```

---

# Generalizability

--

- What if we only used one season?

--

  - Do we think our conclusions would "generalize" (i.e., apply to) other seasons?

--

  - For example, is the turnover-win relationship the same in the 2017 season as the 2018 season?
  
--

  - What about the 2019 season?
  
--

  - Why or why not?
  
--

- Demonstrate using the 2017 data

---

# Generalizability

- Bootstrap + `group_by`

```{r}
bsRes <- NULL

for(i in 1:500) {  # Only 500 simulations this time
  bsRes <- gms %>%
    group_by(yearSeason) %>% #<< Group by the season
    sample_n(size = 100,replace = T) %>% #<< Get 100 observations per season
    group_by(yearSeason,isWin) %>% #<< Then calculate mean tov by season AND win
    summarise(avgTO = mean(tov,na.rm=T),.groups = 'drop') %>%
    ungroup() %>%
    mutate(bsInd = i) %>%
    bind_rows(bsRes)
  
}

```

---

# Plotting the results

```{r}
bsRes %>%
  ggplot(aes(x = avgTO)) + 
  geom_density(alpha = .3)
```

--

- Is this answering our .blue[question]?

---

# Plotting the results

```{r}
bsRes %>%
  ggplot(aes(x = avgTO,fill = isWin)) + 
  geom_density(alpha = .3)
```

- Is this answering our .blue[question]?

---

# Plotting the results

```{r}
bsRes %>%
  ggplot(aes(x = avgTO,fill = isWin)) + 
  geom_density(alpha = .3) + 
  facet_grid(yearSeason~.) #<<
```


---

# Plotting the results

```{r}
p  <- bsRes %>%
  ggplot(aes(x = avgTO,fill = isWin)) + 
  geom_density(alpha = .3) + 
  geom_vline(data = bsRes %>%  #<<
               group_by(yearSeason,isWin) %>%#<<
               summarise(avgTO = mean(avgTO,na.rm=T)),#<<
             aes(xintercept = avgTO,color = isWin),linetype = 'dashed') + #<<
  geom_text(data = bsRes %>%#<<
               group_by(yearSeason,isWin) %>%#<<
               summarise(avgTO = mean(avgTO,na.rm=T)),#<<
             aes(x = avgTO,y = Inf,label = round(avgTO,1)),hjust = 1.1,vjust = 1.1,size = 3,angle = 90) + #<<
  facet_grid(yearSeason~.) 
```

---

# Plotting the results

```{r}
p
```

---

# Plotting the results

```{r}
p  <- bsRes %>%
  ggplot(aes(x = avgTO,fill = isWin)) + 
  geom_density(alpha = .3) + 
  geom_vline(data = bsRes %>% 
               group_by(yearSeason,isWin) %>%
               summarise(avgTO = mean(avgTO,na.rm=T)),
             aes(xintercept = avgTO,color = isWin),linetype = 'dashed') + 
  geom_text(data = bsRes %>%
               group_by(yearSeason,isWin) %>%
               summarise(avgTO = mean(avgTO,na.rm=T)),
             aes(x = avgTO,y = Inf,label = round(avgTO,1)),hjust = 1.1,vjust = 1.1,size = 3,angle = 90) + #<<
  facet_grid(yearSeason~.) 
```

- *Pause*: try changing `hjust` and `vjust` to see how it moves the text on the plot

---

# Summarizing further

--

- We are *actually* interested in whether winning teams turnover the ball less

--

  - .blue[Science]: never forget your theory / hypothesis!
  
--

- So let's actually calculate this!

--

- The `spread` command to create two columns

```{r}
bsRes %>%
  spread(isWin,avgTO,sep = '_') %>%
  mutate(TO_diff = isWin_FALSE - isWin_TRUE)
```

---

# Generalizability

```{r}
bsRes %>%
  spread(isWin,avgTO,sep = '_') %>% 
  mutate(TO_diff = isWin_FALSE - isWin_TRUE) %>%
  ggplot(aes(x = TO_diff,fill = factor(yearSeason))) + 
  geom_density(alpha = .3)
```

---

# Comparing across seasons

```{r}
p <- bsRes %>%
  spread(isWin,avgTO,sep = '_') %>%
  mutate(TO_diff = isWin_FALSE - isWin_TRUE) %>%
  ggplot(aes(x = TO_diff,group = yearSeason)) + 
  geom_density(alpha = .3) + 
  geom_vline(xintercept = 0) + 
  geom_text(data = bsRes %>%
             spread(isWin,avgTO,sep = '_') %>%
             mutate(TO_diff = isWin_FALSE - isWin_TRUE) %>%
             group_by(yearSeason) %>%
             summarise(conf = mean(TO_diff > 0),
                       TO_diff = mean(TO_diff),
                       y = .25),
            aes(x = TO_diff,y = y,label = paste0(round(conf*100,1),'%'))) + 
  facet_grid(yearSeason ~.)
```

---

# Comparing across seasons

```{r}
p
```


---

# Visualization is **DEEP**

```{r}
toplot <- bsRes %>%
  spread(isWin,avgTO,sep = '_') %>%
  mutate(TO_diff = isWin_FALSE - isWin_TRUE)

tmp <- density(toplot$TO_diff)
p <- data.frame(x = tmp$x,y = tmp$y, area = tmp$x >= 0) %>%
  ggplot(aes(x = x,ymin = 0,ymax = y,fill = area)) + 
  geom_ribbon(alpha = .6) + 
  geom_vline(xintercept = 0,linetype = 'dashed',size = 1.1) + 
  annotate(geom = 'text',x = mean(toplot$TO_diff),y = .25,
           label = paste0("Losing team had\nmore turnovers in\n",round(mean(toplot$TO_diff > 0),3)*100,"% of\nBootstraps"), #<<
           hjust = .5) + 
  labs(title = 'Difference in Turnovers by Game Outcome',
       subtitle = '1,000 Bootstrapped Estimates from 2016-2019 Seasons',
       x = 'Losing Team Turnovers minus Winning Team Turnovers',
       y = 'Density of Simulated Games') + 
  scale_fill_manual(name = 'Who Had More Turnovers',
                    values = c('grey60','slateblue'),labels = c('Winning Team','Losing Team')) + 
  theme(panel.background = element_blank(),
        legend.position = 'bottom')
```


---

# Visualization is **DEEP**

```{r,echo=F,fig.width=9,fig.height=6}
p
```

--

- **Side-bar**: If asked in an interview to reproduce this plot, could you? (*Actual* question asked for a research position at UPenn.)


