---
title: "Regression"
subtitle: "Part 1b"
author: "Prof. Weldzius"
institute: "Villanova University"
# date: "Slides Updated: `r Sys.Date()`"
output:
  xaringan::moon_reader:
    # self_contained: true
    chakra: libs/remark-latest.min.js
    lib_dir: libs
    css:
      - default
      - css/lexis.css
      - css/lexis-fonts.css
    #seal: false
    nature:
      highlightStyle: github
      highlightLines: true
      countIncrementalSlides: false
      slideNumberFormat: "%current%"
      ratio: "16:9"

---

```{css,echo = F}
.small .remark-code { /*Change made here*/
  font-size: 85% !important;
}
.tiny .remark-code { /*Change made here*/
  font-size: 50% !important;
}
```

```{r,include=F}
set.seed(123)
options(width=60)
knitr::opts_chunk$set(fig.align='center',fig.width=9,fig.height=5,message=F,warning=F)
def.chunk.hook  <- knitr::knit_hooks$get("chunk")
knitr::knit_hooks$set(chunk = function(x, options) {
  x <- def.chunk.hook(x, options)
  ifelse(options$size != "normalsize", paste0("\n \\", options$size,"\n\n", x, "\n\n \\normalsize"), x)
})
```

# Agenda

1. Introducing the Data

2. Demonstrating Regressions

---

# Regression

--

- Let's demonstrate what we learned last time with the `sc_debt` data

```{r,message = F,warning=F}
require(tidyverse)

debt <- read_rds('../data/sc_debt.Rds')

glimpse(debt)
```

---

# Research Camp

- .blue[Research Question]: What is the relationship between SAT scores and median future earnings?

--

- .blue[Theory]: Students with higher SAT scores work harder and have learned more. Employers reward these attributes with higher wages in the private market.

--

- .blue[Hypothesis]: The relationship between SAT scores and future earnings should be positive.

--

  - **NB**: Important caveats to this simplistic theory!
  
--

  - Socioeconomic status: predicts both higher SAT scores and higher wages
  
--

  - **Correlation $\neq$ Causation**
  
---

# Set Up

- Linking .blue[Theory] to .red[Data]

--

- Our SAT scores are theorized to explain future earnings

--

  - Thus the SAT scores are the independent / explanatory / predictor variable $X$
  
  - And earnings are the dependent / outcome variable $Y$

---

# Regression

- There is a simple recipe to follow

--

- And it is exactly how the syllabus for the class is designed!

--

  1. Look at your data to identify missingness (**Wrangling**)
  
  2. **Univariate** visualization of your variables 
  
  3. **Multivariate** visualization of your variables 
  
  4. **Regression** (today)
  
  5. Evaluation of **errors** (next lecture)

---

# Step 1: Look

- Why worry about **missingness**?

--

1. **Substantive:** external validity

2. **Technical:** cross validation won't work!

--

```{r}
summary(debt %>% select(sat_avg,md_earn_wne_p6))
```

---

# Step 2: Univariate Viz

- Why visualize both $Y$ and $X$?

--

1. **Substantive:** *See* which units you are talking about

2. **Technical:** Adjust for *skew*

---

# Step 2: Univariate Viz

- Why visualize both $Y$ and $X$?

```{r,warning=F,message=F}
debt %>%
  ggplot(aes(x = sat_avg)) + 
  geom_histogram()
```

---

# Step 2: Univariate Viz

- Why visualize both $Y$ and $X$?

```{r,warning=F,message=F}
debt %>%
  ggplot(aes(x = md_earn_wne_p6)) + 
  geom_histogram()
```

---

# Step 3: Multivariate 

- Eyeball the relationship first!

--

```{r,warning=F,message=F}
debt %>%
  ggplot(aes(x = sat_avg,y = md_earn_wne_p6)) + 
  geom_point()
```

---

# Step 3: Multivariate Viz

- Adding regression line

```{r,warning=F,message = F}
debt %>%
  ggplot(aes(x = sat_avg,y = md_earn_wne_p6)) + 
  geom_point() + geom_smooth(method = 'lm',se = F)
```

---

# Step 3: Multivariate Viz

- Let's focus on two schools

```{r,warning=F,message=F}
toplot <- debt %>%
  mutate(hl = ifelse(unitid %in% c(100654,179265),'hl','none')) # Choosing two examples
p2 <- toplot %>%
  ggplot(aes(x = sat_avg, y = md_earn_wne_p6,color = hl,group = 1,alpha = hl)) + 
  geom_point(data = toplot %>% filter(hl == 'none')) + 
  geom_point(data = toplot %>% filter(hl == 'hl'),size =3) + 
  scale_alpha_manual(values = c(1,.3)) + 
  scale_color_manual(values = c('red','black')) + 
  geom_smooth(method = 'lm',se = F) + 
  theme(legend.position = 'none') + 
  labs(title = "Graduate Earnings and SAT Scores",
       subtitle = "By School",
       x = "Average SAT Score",
       y = "Median Earnings of Grads")
```

---

# Step 3: Multivariate Viz

- Adding regression line

```{r,warning=F,message=F}
p2
```

---

# Step 3: Multivariate Viz

- Defining $\varepsilon$

```{r,warning=F,message=F}
p3 <- toplot %>%
  ggplot(aes(x = sat_avg, y = md_earn_wne_p6,color = hl,group = 1,alpha = hl)) + 
  geom_point(data = toplot %>% filter(hl == 'none')) + 
  geom_point(data = toplot %>% filter(hl == 'hl'),size =3) + 
  scale_alpha_manual(values = c(1,.3)) + 
  scale_color_manual(values = c('red','black')) + 
  geom_smooth(method = 'lm',se = F) + 
  annotate(geom = 'segment',
           x = toplot %>% filter(hl == 'hl') %>% .$sat_avg,
           y = toplot %>% filter(hl == 'hl') %>% .$md_earn_wne_p6,
           xend = toplot %>% filter(hl == 'hl') %>% .$sat_avg,
           yend = c(27500,41000),color = 'red',lwd = 1.2) + 
  theme(legend.position = 'none') + 
  labs(title = "Graduate Earnings and SAT Scores",
       subtitle = "By School",
       x = "Average SAT Score",
       y = "Median Earnings of Grads")
```

---

# Step 3: Multivariate Viz

- Measuring errors

```{r,warning=F,message=F}
p3
```
---

# The Data Scientist's Trade-off

--

- Those mistakes seem pretty big!

--

- Why not use a curvier line?

```{r,echo = F}
p4 <- toplot %>%
  ggplot(aes(x = sat_avg, y = md_earn_wne_p6,color = hl,group = 1,alpha = hl)) + 
  geom_point(data = toplot %>% filter(hl == 'none')) + 
  geom_point(data = toplot %>% filter(hl == 'hl'),size =3) + 
  scale_alpha_manual(values = c(1,.3)) + 
  scale_color_manual(values = c('red','black')) + 
  geom_smooth(method = 'gam',se = F) + 
  annotate(geom = 'segment',
           x = toplot %>% filter(hl == 'hl') %>% .$sat_avg,
           y = toplot %>% filter(hl == 'hl') %>% .$md_earn_wne_p6,
           xend = toplot %>% filter(hl == 'hl') %>% .$sat_avg,
           yend = c(27500,41000),color = 'red',lwd = 1.2) + 
  theme(legend.position = 'none') + 
  labs(title = "Graduate Earnings and SAT Scores",
       subtitle = "By School",
       x = "Average SAT Score",
       y = "Median Earnings of Grads")

p5 <- toplot %>%
  ggplot(aes(x = sat_avg, y = md_earn_wne_p6,color = hl,group = 1,alpha = hl)) + 
  geom_point(data = toplot %>% filter(hl == 'none')) + 
  geom_point(data = toplot %>% filter(hl == 'hl'),size =3) + 
  scale_alpha_manual(values = c(1,.3)) + 
  scale_color_manual(values = c('red','black')) + 
  geom_smooth(method = 'loess',se = F,span = .1) + 
  annotate(geom = 'segment',
           x = toplot %>% filter(hl == 'hl') %>% .$sat_avg,
           y = toplot %>% filter(hl == 'hl') %>% .$md_earn_wne_p6,
           xend = toplot %>% filter(hl == 'hl') %>% .$sat_avg,
           yend = c(26500,43000),color = 'red',lwd = 1.2) + 
  theme(legend.position = 'none') + 
  labs(title = "Graduate Earnings and SAT Scores",
       subtitle = "By School",
       x = "Average SAT Score",
       y = "Median Earnings of Grads")

p6 <- toplot %>%
  ggplot(aes(x = sat_avg, y = md_earn_wne_p6,color = hl,group = 1,alpha = hl)) + 
  geom_point(data = toplot %>% filter(hl == 'none')) + 
  geom_point(data = toplot %>% filter(hl == 'hl'),size =3) + 
  scale_alpha_manual(values = c(1,.3)) + 
  scale_color_manual(values = c('red','black')) + 
  geom_smooth(method = 'loess',se = F,span = .025) + 
  annotate(geom = 'segment',
           x = toplot %>% filter(hl == 'hl') %>% .$sat_avg,
           y = toplot %>% filter(hl == 'hl') %>% .$md_earn_wne_p6,
           xend = toplot %>% filter(hl == 'hl') %>% .$sat_avg,
           yend = c(26500,43000),color = 'red',lwd = 1.2) + 
  theme(legend.position = 'none') + 
  labs(title = "Graduate Earnings and SAT Scores",
       subtitle = "By School",
       x = "Average SAT Score",
       y = "Median Earnings of Grads")

p7 <- toplot %>%
  ggplot(aes(x = sat_avg, y = md_earn_wne_p6,color = hl,group = 1,alpha = hl)) + 
  geom_point(data = toplot %>% filter(hl == 'none')) + 
  geom_point(data = toplot %>% filter(hl == 'hl'),size =3) + 
  scale_alpha_manual(values = c(1,.3)) + 
  scale_color_manual(values = c('red','black')) + 
  geom_line(color = 'black') + 
  theme(legend.position = 'none') + 
  labs(title = "Graduate Earnings and SAT Scores",
       subtitle = "By School",
       x = "Average SAT Score",
       y = "Median Earnings of Grads")
```

---

# The Data Scientist's Trade-off

- Those mistakes seem pretty big!

- Why not use a curvier line?

```{r,warning=F,message=F,echo = F}
p4
```

---

# The Data Scientist's Trade-off

- Those mistakes seem pretty big!

- Why not use a curvier line?

```{r,warning=F,message=F,echo = F}
p5
```

---

# The Data Scientist's Trade-off

- Those mistakes seem pretty big!

- Why not use a curvier line?

```{r,warning=F,message=F,echo = F}
p6
```

---

# The Data Scientist's Trade-off

- Those mistakes seem pretty big!

- Why not use a curvier line?

```{r,warning=F,message=F,echo = F}
p7
```

---

# The Data Scientist's Trade-off

--

- Want to **reduce complexity**

--

- But also want to be **accurate**

--

- What is the right answer?

--

  - It depends on your .blue[theory] and the .red[data]
  
--

  - It is context-dependent
  
--

- And this is still only using *linear regression models*!

--

  - This is a deep area of study, for those interested
  

---

# Step 4: Regression

- Introducing the `lm(formula,data)` function

--

- Two inputs to care about:

--

  - `formula`: Code for $Y = \alpha + \beta X$
  
  - `data`: What is the data we are using?
  
- `formula` is written as `Y ~ X`

--

  - `R` will calculate $\alpha$ and $\beta$ for us
  
  - Just need to tell it what is $Y$ (`md_earn_wne_p6`) and $X$ (`sat_avg`)
  
  - The tilde (`~`) is `R`'s version of the equals sign
  
--

- Save the model to an object

```{r}
model_earn_sat <- lm(formula = md_earn_wne_p6 ~ sat_avg,data = debt)
```

---

# Step 4: Interpretation

- What is in this object?

--

- The regression results! Look at them with `summary()`

```{r}
summary(model_earn_sat)
```

---

# Step 4: Interpretation

- Starting with the first column called `Estimate`

--

- 1st Row `(Intercept)` is $\alpha$: the predicted value of $Y$ when $X$ is zero

--

  - Schools with average SAT scores of 0 produce graduates who earn -$12,053.87
  
  - Sensible?
  
--

- 2nd Row `sat_avg` is the $\beta$: the increase in $Y$ when $X$ increases by one

--

  - For each unit increase in the average SAT score, recent graduates earn $42.60 more
  
  - Sensible?
  
---

# Step 4: Interpretation

- Other 3 columns?

--

  - `Std. Error` is the "standard error"
  
  - `t value` is the "t-statistic"
  
  - `Pr(>|t|)` is the "p-value"
  
--

- t-statistic = Estimate / standard error

- p-value = function(t-statistic)

--

  - Only really need to remember the p-value for this course
  
  - This is 1 minus confidence
  
  - The lower the p-value, the more confident we are that the `Estimate` is not zero
  
---

# Step 4: Interpretation

```{r}
summary(model_earn_sat)
```

---

# Another Example

- We will come back to the RMSE next class

--

- For now, let's try with a different research question!

--

- .blue[What is the relationship between admissions and future earnings?]

--

  - .blue[Theory]: More selective schools are more prestigious

  - .blue[Hypothesis]: There should be a negative relationship between the admissions rate and future earnings
  
<!-- --- -->

<!-- # Do It Together! -->

<!-- 1. Look at the data and acknowledge missingness -->

<!-- 2. Univariate visualization of $X$ and $Y$ -->

<!-- 3. Multivariate visualization of $X$ and $Y$ -->

<!-- 4. Regression -->

